---
title: "A1_McLEOD_20294254"
author: "AHM"
date: "2025-01-07"
output: html_document
---
Assignment 1
Aidan McLeod
20294254

Repository link: https://github.com/AidanMcLeod/BIOL432A1

1. Loading the revised measurements csv generated in "dataGenerato.R" and "volumeEstimato.R:
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

#load file
limb_measurements <- read.csv("measurements.csv")
head(limb_measurements)
```

2. Sorting the variables in the file into order:
```{r}
#sort data by species, then by observer, then by limb volume
measurements_sorted <- limb_measurements %>% 
  arrange(species_name, observer_name, Volume)
```

3. Table of the calculated average volumes for each species:
```{r}
#calculate average volume for each species 
avg_volume <- measurements_sorted %>% 
  group_by(species_name) %>% 
  summarize(Average_Volume = mean(Volume, na.rm = TRUE))

#display the average volume table
avg_volume
```
4. Table of the number of observations by observers per species:
```{r}
#Count the number of observations for each combination of species and observer 
observation_count <- measurements_sorted %>%
  group_by(species_name, observer_name) %>% 
  summarize(Num_Observations = n())

#Display the observation count table
observation_count
```

5. Boxplot of the distribution of limb volumes (cm^3) in each species:
```{r}
#Create boxplot comparison of limb volumes between species 
ggplot(measurements_sorted, aes(x = species_name, y = Volume)) + 
  geom_boxplot() + 
  theme_minimal() + 
  labs(title = "Comparison of Limb Volumes Across Species",
       x = "Species",
       y = "Limb Volume (cm^3)") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

6. Histogram panels showing the frequency distribution of limb volumes in each species:
```{r}
#Convert data to long format for easier plotting
measurements_long <- measurements_sorted %>% 
  gather(key = "Measurement_Type", value = "Value", Volume)

#Create multi-panel plot histograms of limb volume by species
ggplot(measurements_long, aes(x = Value)) +
  geom_histogram(bins = 15, fill = "skyblue", color = "black", alpha = 0.7) +
  facet_wrap(~ species_name, scales = "free_x") +
  theme_minimal() +
  labs(title = "Distribution of Limb Volume by Species",
       x = "Limb Volume (cm³)",
       y = "Frequency")
```
